---
title: "Cervical cancer: APAC example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cervical cancer: APAC example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      comment = "#>",
                      eval = FALSE)
library(tidyverse)

# To run another analysis change "eval = FALSE" in line 14 to "eval = TRUE" e follow all steps bellow.
```

## 1. About this Vignette

This vignette shows how to use the `OncoDatasusR` package to analyze cervical cancer data. In this example, we will download and process data from High Cost Procedures Authorizations 
(APAC in Portuguese) of the Outpatient Information System (SIA) to count the number of patients with cervical cancer starting treatment in the Brazilian public health system.

This vignette was written to allow the reader to replicate the Cervical Example (paper in *peer review*) without having to download data from DATASUS and to obtain identical results. 
To do this, it is necessary to load the RData file available in [GitHub](https://github.com/PatriciaKZ/OncoDatasusR/tree/main/vignettes/exported_data). 

```{r, include=FALSE, eval = TRUE}
load("exported_data/apac_cervical_analysis.RData")
```

## 2. Introduction to OncoDatasusR

The `OncoDatasusR` package provides functions to download and process data from the Department of Informatics of the Unified Health System (DATASUS), which is the main 
public health database in Brazil. Along with the process of creating a dataset with unique patients, this package also provides some information to evaluate data quality.
Data quality evaluation is a important step since the proposed patient counting methodology may face inaccuracies due to data limitations 
and the deduplication process. Despite our attempts to reduce the number of patients counted more than once, issues such as incomplete or incorrect records may influence 
results. A detailed step-by-step analysis of the creation of the final dataset is essential to better understand the limitations and support accurate interpretations of results.

## 3. Prepare the environment

### Load required packages

```{r setup}
library(OncoDatasusR)
library(tidyverse)
library(janitor)
```

## 4. Download Data from DATASUS

The first step involves extracting data directly from the DATASUS website using 
File Transfer Protocol (ftp://ftp.datasus.gov.br/dissemin/publicos/SIASUS/200801_/Dados/).
This will be done using the get_datasus function which uses the functions detailed below.

+ retrieve_filelist: search in the DATASUS ftp all the desired files and years 
of interest that are available for download.

+ download_datasus: extract the files by Federation Unit/year of interest and save 
them in the ‘datasus_files’ folder.

+ fetch_files: creates a list with the names of the ‘.csv files’ saved in the 
folder ‘datasus_files’

+ merge_datasus: merges all .csv files into a single file. 

### 4.1. Download data of Chemotherapy (AQ) APACs

In the function below we are specifying that we want to extract data from patients 
with cervical cancer "C53", from "all" federation units in Brazil, that 
the name of the data files to be saved will start with "cervical", which will be 
the "AQ" data processed between the years from "2008" and "2023" and which will be saved 
in a folder called 'datasus_files' created in the current working directory. 
If you want to choose another name on the folder, replace NULL with the desired name.

```{r download_data_aq}
get_datasus(
  icd_list = "C53",
  fu_list = "all",
  export_file_code = "cervical",
  database = "AQ",
  first_year = 2008,
  last_year = 2023,
  output_dir = NULL
)
#>  ...
#>  trying URL 'ftp://ftp.datasus.gov.br/dissemin/publicos/SIASUS/200801_/Dados/AQTO2312.dbc'
#>  Content type 'unknown' length 4541 bytes
#>  Reading: datasus_files/AQ_AC_cervical_2008_2023.csv
#>  ...
#>  Writing: datasus_files/AQ_cervical_2008_2023_full.csv
#>  File saved successfully.
```

### 4.2. Download data of Radiotherapy (AR) APACs

In the function below we are repeating the extraction process, now choosing "AR" data.

```{r download_data_ar}
get_datasus(
  icd_list = "C53",
  fu_list = "all",
  export_file_code = "cervical",
  database = "AR",
  first_year = 2008,
  last_year = 2023,
  output_dir = NULL
)
#>  ...
#>  trying URL 'ftp://ftp.datasus.gov.br/dissemin/publicos/SIASUS/200801_/Dados/ARTO2312.dbc'
#>  Content type 'unknown' length 4541 bytes
#>  Reading: datasus_files/AR_AC_cervical_2008_2009.csv
#>  ...
#>  Writing: datasus_files/AR_cervical_2008_2023_full.csv
#>  File saved successfully.
```

## 5. SIA-SUS Data Preparation

The second step involves the creation of a database with unique patients. It will be done using the extracted data and following rules established for this purpose.
Quality assessment is available to support subsequent analyses.

### 5.1. Read APAC files

The function below reads all AQ and AR files that were extracted creating an object with all extracted APACs. When dir is NULL, the function will look for the files 
in the "datasus_files" folder in the current working directory. Otherwise, the function will look for the files in the directory defined by 'dir' argument.
Only the variables of interest chosen from those available will be kept in this object. By default de function will select 15 variables.

* database (AQ or AR), 
* cod_paciente (encrypted code of the National Health Card, CNS: Cartão Nacional de Saúde), 
* cid_prim (primary ICD with 3 digits), 
* CID_completo (complete ICD with four digits), 
* proc (code of the procedure performed), 
* data_diagnostico (date of diagnosis), 
* ano_diagnostico (year of diagnosis), 
* data_atendimento, 
* ano_atendimento, 
* cep_paciente (zip code of the patient’s address), 
* sexo (sex), 
* idade (age at diagnosis with 2 digits according to the DATASUS age code),  
* cod_idade (age code, as defined by the DATASUS), 
* estadi (staging),
* ap_munpcn (code of the city of the residence of the patient). 

By including the ‘select_variables’ argument it is possible to add other variables among those available. In this example we chose the variable AP_MVM (month and year of APAC 
processing) as an additional variable.

The 'group_icd' argument can be used to group ICD codes for future analysis. If it is used, it is necessary to inform the icd_group_file_path. Be aware that the APACS 
databases (AQ and AR) are subjected to updates of the previous competence (see PAPER in *peer review*).

The 'med_proced_to_remove' argument can be used to exclude procedures of non-interest.

The output is a list with:

* The APAC dataframe (where each line will display information referring to one APAC),
* The diag_list list that will have some consistency check information (see PAPER).

```{r read_files}
apac <- read_apac(
  read_bases = c("AQ", "AR"),
  aq_file = "AQ_cervical_2008_2023_full.csv",
  ar_file = "AR_cervical_2008_2023_full.csv",
  dir = NULL,
  first_year = 2008,
  last_year = 2023,
  file_code = "cervical",
  run_diagnostics = TRUE,
  group_icd = FALSE,
  export_procedure_table = TRUE,
  select_variables = "AP_MVM"
)
#>  Reading: datasus_files/AQ_cervical_2008_2023_full.csv
#>  Reading: datasus_files/AR_cervical_2008_2023_full.csv
#>  Writing: datasus_files/cervical_2008_2023_procedures.csv

str(apac, max.level = 1)
#List of 2
# $ APAC     :'data.frame':	1119834 obs. of  16 variables:
# $ diag_list:List of 7```
```

#### 5.1.1. Quality evaluation

The total number of APACs in the dataset is:

```{r n_apac}
n_apac_total <- nrow(apac$APAC)
```

```{r p_n_apac , eval=TRUE}	
n_apac_total
```

The table bellow shows the number of APACs by ICD and SIA database. It is useful for understanding the distribution of APACs by ICD and SIA database. The table also 
provides a description of the ICD codes, by joining the table with the icd_description dataset from the package, to facilitate interpretation. 

```{r table_icd_sia}
table_icd_sia <- apac$diag_list$ICD_base_table %>%
  rbind(colnames(apac$diag_list$ICD_base_table)) %>%
  select(-Total) %>%
  t() %>%
  data.frame() %>%
  row_to_names(1) %>%
  left_join(OncoDatasusR::icd_description, by = c(Base = "cd_cid")) %>%
  replace_na(list(group = " ", ds_cid = " ")) %>%
  relocate(Base, ds_cid, AQ, AR, Total)
```
```{r p_table_icd_sia , eval=TRUE}	
table_icd_sia
```

The table bellow shows the number of unique CNS numbers. It is useful for identifying the number of unique patients and for understanding the distribution of patients by CNS 
and SIA database.

```{r table_cns}
table_cns_sia <- apac$diag_list$unique_patient_code_all
```
```{r p_table_cns, eval=TRUE}	
table_cns_sia
```

The table bellow shows the data completeness and accuracy of the APAC dataset created. It is useful for identifying the number of records with missing or inconsistent data and 
for understanding the quality of the data.

```{r table_data_quality}
colnames(apac$diag_list$error_table_all) <- c(
  "Procedure",
  "Date of Diagnosis",
  "Date of Attendance",
  "Patient ZIP code",
  "Sex",
  "Age",
  "Age Code",
  "Staging",
  "Patient Municipality Code",
  "Date of Processing"
)

table_data_quality <- apac$diag_list$error_table_all %>% t()
```
```{r p_table_data_quality , eval=TRUE}	
table_data_quality
```

The table bellow shows the number of APACs by year of diagnosis. It is useful for identifying the distribution of APACs by year and for understanding the trends 
in the number of APACs over time and the inconsistencies in the year of diagnosis and year of procedure data.

```{r table_year}
table_year <- apac$diag_list$dates_year_table %>%
    replace_na(list(ano_atendimento_registros = 0))
```
```{r p_table_year , eval=TRUE}
head(table_year)
```

The table bellow shows the number of APACs by ICD category. It is useful for understanding the distribution of APACs by ICD.

```{r table_icd}
table_icd <- apac$diag_list$icd_ds_tbl
```
```{r p_table_icd, eval=TRUE}	
table_icd
```

The table bellow shows the number of APACs by procedure type. In this example we decided not to exclude any procedure from the analyses. However, it is recommended 
that procedures that are not appropriate for the cancer under study be excluded due to potential errors.

```{r table_procedure}
table_proced_total <- apac$diag_list$procedure_table
```

```{r p_table_procedure, eval=TRUE}	
head(table_proced_total,5)
```

The table bellow shows the number of records before and after applying a filter to remove APACs with non-interested procedures. 
If a vector with list of procedures was not provided in the med_proced_to_remove argument of the read_apac function the dif column will be zero.

```{r proc_table}	
table_procedures <- apac$diag_list$proc_diag_table
```

```{r p_proc_table , eval=TRUE}
table_procedures
```

### 5.2. Unify Records with the same CNS Number

The function below summarizes the data extracted by the CNS (National Insurance Card). This will be done by selecting the APAC with the earliest diagnosis 
and procedure date.
If the year of the earliest date of diagnosis is before 1998 (10 years before 2008), the second oldest date will be considered the diagnosis date. 
Finally, when the second oldest date is before 1998 or the oldest date is after 2024, the CNS is excluded from the data set. This function also creates the new 
variables listed bellow.

* idade_anos: patient´s age on the day of the procedure calculated using the variables idade and cod_idade.
* idade_diagnostico: patient´s age at diagnosis calculated using the age on the day of the procedure and the date of diagnosis. 
* tratamento_ano: procedures performed during the study period.
* other variables for data quality

The max_age argument determines the maximum age of patients that will be kept in the dataset.The age considered in this filter can be the age at diagnosis 
(age_diagnosis) or the age on the day of the procedure (age_attendance).

The output is a list with:

* The APAC tibble (where each line will display information referring to one CNS),
* Seven other tibbles with quality information (see PAPER).

```{r summarise_data}
apac_sum <- summarise_apac(apac$APAC, run_diagnostics = TRUE)

str(apac_sum)
#> List of 8
#>  $ APAC                        : tibble [191,048 × 27] (S3: tbl_df/tbl/data.frame)
#>  $ apac_per_patient            : tibble [67 × 2] (S3: tbl_df/tbl/data.frame)
#>  $ dt_table_diag               : tibble [5 × 5] (S3: tbl_df/tbl/data.frame)
#>  $ dt_diag_unique_patient      : tibble [1 × 1] (S3: tbl_df/tbl/data.frame)
#>  $ dt_diag_unique_apac         : tibble [1 × 1] (S3: tbl_df/tbl/data.frame)
#>  $ dt_diag_error_tbl           : tibble [3 × 3] (S3: tbl_df/tbl/data.frame)
```

#### 5.2.1. Quality evaluation

The table bellow shows the number of CNS codes by the number of different diagnosis date. We can see that 129,564 (67.8%) CNS codes have just one diagnosis date.

```{r table_inconsist_date}	
table_inconsist_date <- apac_sum$dt_table_diag %>%
  mutate(n_dt_diag = ifelse(n_dt_diag == "5+", "5 or more", n_dt_diag))
```

```{r p_table_inconsist_date , eval=TRUE}
table_inconsist_date
```

The table below shows the number of CNS codes with coherent diagnosis dates (190,370), non-coherent ones that were recovered using the second-smallest date rule (322), 
and inconsistent ones that were not recovered (319).


```{r table_date}	
total_dt_error <- apac_sum$dt_diag_error_tbl
```

```{r p_table_date , eval=TRUE}
total_dt_error
```

The table bellow shows the number of CNS code by the number of APACs. We can see that the most frequent number of APACs is 5.

```{r table_apac_cns}
table_apac_cns <- apac_sum$apac_per_patient
```

```{r p_table_apac_cns , eval=TRUE}
table_apac_cns
```

The number of CNS codes by the number of APACs can be visualized in the plot bellow.

```{r plot_cns_pcn}	
plot_cns_pcn <- apac_sum$apac_per_patient %>%
  mutate(p = cumsum(n) / sum(n)) %>%
  filter(p < 0.99) %>%
  rename(`APAC by patient` = linhas, `Patient (CNS)` = n) %>%
  ggplot(aes(x = `APAC by patient`, y = `Patient (CNS)`)) +
  geom_col(fill = "steelblue") +
  theme_minimal() +
  theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    axis.line = element_line(color = "black")
  ) +
  xlab("") +
  labs(caption = "Obs: the x axis includes up to 99% of the most frequent APACs, and the most extreme
        values were omitted from the visualization")
```

```{r p_plot_cns_pcn , eval=TRUE, fig.width = 7, fig.height = 3.5, fig.align = "center"}	
plot_cns_pcn
```

### 5.3. Unify Records from Patients with Two or More CNS Numbers

On this step the data is prepared to search for patients with two or more different CNS numbers and combine the information from the identified cases.

#### 5.3.1. Prepare APAC Data 

To analyze duplicates, the function below, among other things, creates a key variable for each CNS code, taking into account the patient's ICD, gender, zip code (only the first five digits) 
and date of diagnosis. The comparison of patients also takes into account the difference in year of birth calculated by plus or minus one year. Finally, each row in the table receives an 
identification number (id column):

```{r prepare_linkage}
apac_linkage <- prepare_linkage(apac_sum$APAC, run_diagnostics = TRUE)

str(apac_linkage, max.level = 1)
#>  List of 2
#>   $ APAC_table      : tibble [191,048 × 30] (S3: tbl_df/tbl/data.frame)
#>   $ key_var_NA_table: tibble [1 × 3] (S3: tbl_df/tbl/data.frame)

apac_linkage$APAC_table %>% select(id, cod_paciente, chave1, ano_nasc_calc) %>% head(9) 
#> A tibble: 9 × 4
#>     id cod_paciente                     chave1               ano_nasc_calc
#>  <int> <chr>                            <chr>                        <dbl>
#> 1     1 0c34cf680d4a2065abf061e36fc5da1f C53_F_66000_20071112          1949
#> 2     2 272efd08a31f7a3c274f50e531928483 C53_F_66000_20071112          1948
#> 3     3 37ef8603b831867afb4879f1636ea139 C53_F_66000_20071112          1953
#> 4     4 5483d06863da65bcf43384c2fe7c3e40 C53_F_66000_20071112          1957
#> 5     5 9caf771ad914c2a1d2d58342d0c381ad C53_F_66000_20071112          1950
#> 6     6 b7f7a153ef427b86297487c94e90db57 C53_F_66000_20071112          1975
#> 7     7 ba3ddc86dd95f4bb4d376d057891573e C53_F_66000_20071112          1975
#> 8     8 c53edba550c568c3605989d253c93740 C53_F_66000_20071112          1944
#> 9     9 ec0935c6422469fedcb917f49bab12ec C53_F_66000_20071112          1957
```

#### 5.3.2. Prepare APAC Data: quality information

The table bellow shows the completeness of the key variables. It is useful for identifying the number of records with missing key variables for the deduplication process and for 
understanding the quality of the data.

```{r table_key_var}	
table_key_var <- apac_linkage$key_var_NA_table
```

```{r p_table_key_var , eval=TRUE}
table_key_var
```

#### 5.3.3. Create Correspondence Table

The function below groups the CNS that contain the same value for the key variable and the same year of birth (allowing a margin of error of +- 1 year).

The output is a table with the potential duplicates. Each row in the table is a group of CNS. Note that, due to the margin of error in the year of birth, a CNS can belong to more 
than one group. The message "Warning message:" means that at least one CNS is in more than one group.

```{r create_correspondence_table}
apac_match_list <- get_match_list(apac_linkage$APAC)
#> Processing time: 37.5572071075439
#> Warning message:
#> In get_match_list(apac_linkage$APAC) :
#> Record appears in more than one group. Verify manually
```

Bellow we can see the last lines of the created table.

* The line 4024 show that ids 1, 2 and 5 are duplicates. 
* The line 4023 show that ids 1 and 2 are duplicates.
* The line 4022 show that ids 1 and 5 are duplicates.

One can see that id=1 belongs to those three groups. When it happens we have to perform some extra code lines.  

```{r show_table}
apac_match_list[nrow(apac_match_list):(nrow(apac_match_list)-6),]
#>     X1 X2 X3 X4 X5
#> 4024  1  2  5 NA NA
#> 4023  1  2 NA NA NA
#> 4022  1  5 NA NA NA
#> 4021  6  7 NA NA NA
#> 4020  4  9 NA NA NA
#> 4019 11 14 16 NA NA
#> 4018 18 19 NA NA NA
```

Checking the values of lines with **id** equal to 1, 2 and 5.

```{r show_records}
apac_linkage$APAC_table %>% select(id, cod_paciente, chave1, ano_nasc_calc) %>% slice(c(1,2,5))
#> A tibble: 3 × 4
#>     id cod_paciente                     chave1               ano_nasc_calc
#>   <int> <chr>                            <chr>                        <dbl>
#> 1     1 0c34cf680d4a2065abf061e36fc5da1f C53_F_66000_20071112          1949
#> 2     2 272efd08a31f7a3c274f50e531928483 C53_F_66000_20071112          1948
#> 3     5 9caf771ad914c2a1d2d58342d0c381ad C53_F_66000_20071112          1950
```

#### 5.3.4. Extra Code 

The first step is to identify the **ids** that belongs to more than one group which, in this example, is a total of 24.

```{r id_multiplicity}
repeated_ids = data.frame(ids = unlist(apac_match_list)) %>%
  count(ids) %>%
  drop_na(ids) %>%
  filter(n > 1) %>%
  select(ids)
repeated_ids_df <- data.frame()

for (rep_id in 1:length(repeated_ids$ids)) {
  print(rep_id)
  repeated_ids_position = which(apac_match_list == repeated_ids$id[rep_id], arr.ind =
                                  TRUE) %>%
    data.frame()
  
  repeated_ids_df = rbind(repeated_ids_df, repeated_ids_position)
}

#> [1] 1
#> [1] 2
#> [1] 3
#> [1] 4
#> [1] 5
#> [1] 6
#> [1] 7
#> [1] 8
#> [1] 9
#> [1] 10
#> [1] 11
#> [1] 12
#> [1] 13
#> [1] 14
#> [1] 15
#> [1] 16
#> [1] 17
#> [1] 18
#> [1] 19
#> [1] 20
#> [1] 21
#> [1] 22
#> [1] 23
#> [1] 24
```

In the table below it is possible to see the groups where the **ids** with multiplicity are found. Observe the pattern of the groups. In the first three lines, we can see that the records
(1, 5), (1, 2), and (1, 2, 5) were grouped. Next, we have the records (18,19,22), (18,19), and (19,22). This occurs due to the way the comparison code for the get_match_list() function 
was written. At this point we have to keep just the lines with more **ids**. For example,we have to exclude lines 4022 and 4023 and keep line 4024.To understand the rationale of this 
step (see PAPER).

```{r id_multiplicity_table}
apac_match_list[unique(repeated_ids_df$row),]
#>      X1  X2  X3  X4  X5
#> 4022   1   5  NA  NA  NA
#> 4023   1   2  NA  NA  NA
#> 4024   1   2   5  NA  NA
#> 4017  18  19  22  NA  NA
#> 4018  18  19  NA  NA  NA
#> 4016  19  22  NA  NA  NA
#> 4003  61  63  64  NA  NA
#> 4004  61  63  64  65  NA
#> 4002  63  65  NA  NA  NA
#> 3988 109 110 111 112 113
#> 3989 109 112 113  NA  NA
#> 3990 109 110 111 113  NA
#> 3887 551 553  NA  NA  NA
#> 3888 551 552  NA  NA  NA
#> 3889 551 552 553  NA  NA
#> 3852 728 729 730  NA  NA
#> 3854 728 730  NA  NA  NA
#> 3853 729 730  NA  NA  NA
#> 3848 740 741 742  NA  NA
#> 3849 740 741  NA  NA  NA
#> 3847 741 742  NA  NA  NA
```

The code bellow will creat two more variables:

* na_n = number of missings in the line
* row_n = numero da linha

```{r new_variables}
match_list_rep <- apac_match_list[unique(repeated_ids_df$row), ]
match_list_rep <- match_list_rep %>%
  mutate(na_n = apply(match_list_rep, 1, function(x)
    sum(is.na(x))),
    row_n = rownames(.))
match_list_rep
#>       X1  X2  X3  X4  X5 na_n row_n
#> 4022   1   5  NA  NA  NA    3  4022
#> 4023   1   2  NA  NA  NA    3  4023
#> 4024   1   2   5  NA  NA    2  4024
#> 4017  18  19  22  NA  NA    2  4017
#> 4018  18  19  NA  NA  NA    3  4018
#> 4016  19  22  NA  NA  NA    3  4016
#> 4003  61  63  64  NA  NA    2  4003
#> 4004  61  63  64  65  NA    1  4004
#> 4002  63  65  NA  NA  NA    3  4002
#> 3988 109 110 111 112 113    0  3988
#> 3989 109 112 113  NA  NA    2  3989
#> 3990 109 110 111 113  NA    1  3990
#> 3887 551 553  NA  NA  NA    3  3887
#> 3888 551 552  NA  NA  NA    3  3888
#> 3889 551 552 553  NA  NA    2  3889
#> 3852 728 729 730  NA  NA    2  3852
#> 3854 728 730  NA  NA  NA    3  3854
#> 3853 729 730  NA  NA  NA    3  3853
#> 3848 740 741 742  NA  NA    2  3848
#> 3849 740 741  NA  NA  NA    3  3849
#> 3847 741 742  NA  NA  NA    3  3847
```

Below is the part that configures the manual evaluation. A new variable should be created indicating the group to which each line belongs. This **MUST BE MODIFIED** as per the example. 
In this example, 7 groups were identified. Note that they will not always be groups of 3 lines as happened here. This is why it is necessary to create the group variable manually.

```{r creat_groups}
match_list_rep$group = factor(c(
  rep(1, 3),
  rep(2, 3),
  rep(3, 3),
  rep(4, 3),
  rep(5, 3),
  rep(6, 3),
  rep(7, 3)
))
match_list_rep
#>  X1  X2  X3  X4  X5 group
#> 4022   1   5  NA  NA  NA     1
#> 4023   1   2  NA  NA  NA     1
#> 4024   1   2   5  NA  NA     1
#> 4017  18  19  22  NA  NA     2
#> 4018  18  19  NA  NA  NA     2
#> 4016  19  22  NA  NA  NA     2
#> 4003  61  63  64  NA  NA     3
#> 4004  61  63  64  65  NA     3
#> 4002  63  65  NA  NA  NA     3
#> 3988 109 110 111 112 113     4
#> 3989 109 112 113  NA  NA     4
#> 3990 109 110 111 113  NA     4
#> 3887 551 553  NA  NA  NA     5
#> 3888 551 552  NA  NA  NA     5
#> 3889 551 552 553  NA  NA     5
#> 3852 728 729 730  NA  NA     6
#> 3854 728 730  NA  NA  NA     6
#> 3853 729 730  NA  NA  NA     6
#> 3848 740 741 742  NA  NA     7
#> 3849 740 741  NA  NA  NA     7
#> 3847 741 742  NA  NA  NA     7
```

The function bellow will identify the lines to be removed.

```{r lines_to_remove}
match_list_rows_to_remove <- match_list_rep %>%
  group_by(group) %>%
  arrange(na_n, .by_group = TRUE) %>%
  mutate(order = 1:n()) %>%
  filter(order != 1) %>%
  ungroup() %>%
  select(row_n) %>%
  mutate(row_n = as.numeric(row_n)) 

str(match_list_rows_to_remove)
#> tibble [14 × 1] (S3: tbl_df/tbl/data.frame)
#>  $ row_n: num [1:14] 3990 4003 3989 4022 4023 ...
```

Remove the lines identified.

```{r removig_lines}
apac_match_list_final <- apac_match_list[-match_list_rows_to_remove$row_n, ]  
```

Look the last lines of the "apac_match_list_final".

```{r showing_lines}
apac_match_list_final[nrow(apac_match_list_final):(nrow(apac_match_list_final) -
                                                     4), ]
#>      X1 X2 X3 X4 X5
#> 4024  1  2  5 NA NA
#> 4021  6  7 NA NA NA
#> 4020  4  9 NA NA NA
#> 4019 11 14 16 NA NA
#> 4017 18 19 22 NA NA
```

#### 5.3.5. Unify information of potential CNS duplicates

The function bellow unifies the information of potential CNS duplicates identified. For each CNS group, only the CNS with the lowest diagnosis/procedure date will be kept.

The output is a list with:
* The APAC tibble (where each line will display information referring to one CNS),
* The cns_count tibble (showing the number of CNSs grouped into groups of 1, 2, 3, ...)

```{r unify_records}
apac_merged <- merge_cns(apac_linkage$APAC, apac_match_list, run_diagnostics = TRUE)

str(apac_merged, max.level = 1)
#>  List of 2
#>   $ APAC     : tibble [186,956 × 33] (S3: tbl_df/tbl/data.frame)
#>   $ cns_count: tibble [3 × 2] (S3: tbl_df/tbl/data.frame) 
```

#### 5.3.6. Unify information of potential CNS duplicates: quality information

The table bellow shows the number of duplicates CNS numbers by patient. 

```{r table_dup_cns}
table_dup_cns <- apac_merged$cns_count
```

```{r p_table_dup_cns , eval=TRUE}
table_dup_cns
```

### 5.4. Filter APAC Data and save the final dataset

The function below filters the data by year of diagnosis and patient age. In the argument "age_type" is possible to choose between the "age_diagnosis" or "age_attendance". Removes patients with inconsistent dates and saves the final dataset in the exported_data folder created in the current directory.

The output is a list with:
* The APAC_table tibble (where each line will display information referring to 
one unique patient),
* The diag_list dataframe with the number of records before and after filtering.

```{r filter_data}
apac_final <- prepare_apac(
  apac_table = apac_merged$APAC,
  first_year = 2008,
  last_year = 2022,
  age_type = "age_diagnosis",
  age_group = "adult",
  sex = "F",
  run_diagnostics = TRUE,
  write = TRUE,
  output_dir = NULL,
  filename = "Base_APAC_cervical_2008_2022_final.csv"
)
#> Writing: exported_data/Base_APAC_cervical_2008_2022_final.csv

str(apac_final, max.level = 1)
#>  List of 2
#>   $ APAC_table: tibble [168,420 × 33] (S3: tbl_df/tbl/data.frame)
#>   $ diag_list :'data.frame':     1 obs. of  17 variables:
```

### 5.5. Numbers of patients step by step

The codes below show the number of patients at each step of the data preparation process.

##### Number of CNS before deduplication

```{r n_before_deplication}
n_before_deduplication <- nrow(apac_linkage$APAC_table)
```
```{r p_n_before_deplication , eval=TRUE}
n_before_deduplication
```

###### Number of CNS after deduplication

```{r n_after_deplication}
n_after_deduplication <- nrow(apac_merged$APAC)
```
```{r p_n_after_deplication , eval=TRUE}
n_after_deduplication
```

###### Number of CNS removed by deduplication

```{r n_unique_patients}
diff_linkage <- nrow(apac_linkage$APAC_table) - nrow(apac_merged$APAC)
```
```{r p_n_unique_patients , eval=TRUE}
diff_linkage
```

###### Percentage of CNS removed by deduplication

```{r var_linkage}
var_linkage <- round(
    (nrow(apac_linkage$APAC_table) -
        nrow(apac_merged$APAC)) / nrow(apac_linkage$APAC_table) *
        100, 2
)
```
```{r p_var_linkage , eval=TRUE}
var_linkage
```

###### Number of patients with inconsistent dates

```{r n_inconsistent_date}
dt_attend_before <- apac_final$diag_list$dt_error_n
```
```{r p_n_inconsistent_date , eval=TRUE}
dt_attend_before
```

###### Number of patients removed by filtering based on the year of diagnosis

```{r n_after_year}
n_before_min_year <- apac_final$diag_list$lt_first_year_n
n_after_max_year <- apac_final$diag_list$gt_last_year_n
```

```{r p_n_after_year , eval=TRUE}
n_before_min_year
n_after_max_year
```

##### Number of patients removed by filtering based on age

```{r n_after_age}
n_lt_min_age <- apac_final$diag_list$lt_min_age
n_gt_max_age <- apac_final$diag_list$gt_max_age
```

```{r p_n_after_age , eval=TRUE}
n_lt_min_age
n_gt_max_age
```

##### Number of patients removed by filtering based on sex

```{r n_after_sex}
n_rm_sex <- apac_final$diag_list$rm_sex
```
```{r p_n_after_sex, eval=TRUE}
n_rm_sex
```

##### Final number of patients

```{r n_final_patients}
n_patients_final <- apac_final$diag_list$final_patient_n
```
```{r p_n_final_patients , eval=TRUE}
n_patients_final
```

## 6. Data Analysis

The functions bellow were created to generate descriptive tables and plots to analyze the data. For suggestion of how to format and present these data in a report, see the 
vignette "cervical_cancer_apac_report.Rmd" in the documentation folder of the package.

### 6.1. Descriptive table

The function below creates a table with the mean age and the number of patients with cervical cancer by sex, group age, geographic region of patient and clinical stage.

The age_group argument allows to choose age group and define the minimum and maximum age filters to be applied in the descriptive analysis. Be sure to select an age_group 
consistent with the one chosen in the prepare_apac function. The split parameter allows to choose the type of split to be used in the analysis.

```{r descriptive_table}
patient_table <- prepare_patient_table(
  apac_final$APAC_table,
  method = "APAC",
  age_group = "adult",
  age_type = "age_diagnosis"
)
```

```{r p_patient_table , eval=TRUE}
patient_table
```

### 6.2. Plot of number of patients with cervical cancer by year of diagnosis

The number of patients with cervical cancer by treatment and year of diagnosis can be visualized using the function below. The y_type argument allows to choose whether to include 
the "number" or the "proportion" of patients in the plot. 


```{r count_patients}
plot_treat <- plot_patient_count(
  apac_final$APAC_table,
  method = "APAC",
  y_type = "number",
  group = "procedure",
  age_group = "adult",
  age_type = "age_diagnosis",
  legend_title = "Treatment",
  include_label = FALSE
) +
  aes(customdata = tratamento)
```

```{r plot , eval=TRUE, fig.width = 7, fig.height = 3.5, fig.align = "center"}	
plot_treat
```

## 7. Saving objects

The code bellow saves the objects created in this vignette to an .RData file. The file is saved in the "exported_data" folder in the working directory. The objects can be loaded 
later for further analysis or to generate the automatic report. 

```{r save_objects}
save(
  patient_table,
  plot_treat,
  n_apac_total,
  table_icd,
  table_procedures,
  table_icd_sia,
  table_cns_sia,
  table_data_quality,
  table_year,
  total_dt_error,
  total_replaced_diag_dt,
  table_proced_total,
  plot_cns_pcn,
  table_apac_cns,
  table_inconsist_date,
  table_key_var,
  table_dup_cns,
  n_before_deduplication,
  n_after_deduplication,
  diff_linkage,
  var_linkage,
  n_before_min_year,
  n_after_max_year,
  dt_attend_before,
  n_lt_min_age,
  n_gt_max_age,
  n_rm_sex,
  n_patients_final,
  file = "exported_data/apac_cervical_analysis.RData"
)
```
