---
title: "Report on the Count of Patients Treated for Pancreatic Cancer"
subtitle: "SIASUS/DATASUS"
output:
  html_document:
    toc: true
    toc_float: true
vignette: >
  %\VignetteIndexEntry{Report on the Count of Patients Treated for Pancreatic Cancer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(OncoDatasusR)
library(kableExtra)
library(DT)
library(janitor)
library(tidyverse)
library(plotly)

load("exported_data/apac_pancreas_analysis.RData")
# Check the directory where the file 'exported_data/apac_pancreas_analysis.RData' is saved and adjust the path accordingly.
```

# 1. Introduction 

This report presents the results of the analysis of the count of patients treated with chemotherapy, hormone therapy, immunotherapy and/or radiotherapy for pancreatic cancer, based on data provided by DATASUS and using the methodology proposed by Ziegelmann et all (paper in *peer review*) and implemented by the `OncoDatasusR` package. The R code used to generate the results presented here is available in the package documentation (see vignette "pancreatic_cancer_apac_example").

The following criteria were used to select the records (patients) for the analysis:

* Data source: Outpatient Information System  (SIASUS);
* Date of data extraction: 2024-11-28;
* Period of data extraction: Records processed between 2008 to 20203;
* Period of analyses: patients diagnosed between 2008 to 2022;
* ICD code: C25  Malignant neoplasm of pancreas.

This report is divided into two parts:

* Part 1: presents the number of unique patients identified as well as descriptive tables and graphs.
* Part 2: presents the step-by-step process for creating the unique patient dataset used in part 1 along with results for evaluating the quality of the created dataset.

# 2. Part 1: Number of Unique Patients

The number of patients starting treatment counted was **`r n_patients_final`**. The treatment considered were all chemotherapy, hormone therapy, immunotherapy and radiotherapy procedures available in SUS. The characteristics of the patients are described in the table bellow.

## 2.1. Descriptive Table of Pancreatic Cancer Patients

```{r descriptive_table, echo = FALSE}
patient_table %>%
  select(-row_names) %>%
  kable(align = c("l", "c"), col.names = c("", paste0("n = ", n_patients_final))) %>%  # add n patients to the 'Total' column
        add_header_above(c("Variable", "Total"),
                         bold = TRUE,
                         align = c("l", "c")) %>%
          kable_classic() %>%
          kableExtra::group_rows("Age Years", 1, 5) %>%
          kableExtra::group_rows("Sex", 6, 7) %>%
          kableExtra::group_rows("Region", 8, 12) %>%
          kableExtra::group_rows("Clinical Stage", 13, 18) %>%
          kable_paper(full_width = F, "hover") %>%
          add_footnote("NI: Not informed", notation = "none")
```

## 2.2. Distribution of Patients by Treatment Type and Year of Diagnosis

The figure below presents the distribution of patients by treatment type and year of diagnosis, considering all types of outpatient treatments received by each patient in the whole period of analysis.

```{r , fig.width=8, echo = FALSE}
ggplotly(plot_treat, tooltip = c("x", "y", "customdata")) %>%
  style(
    hovertemplate = paste(
      "<b>Year:</b> %{x}<br>",
      "<b>N of patients:</b> %{y:.0f}<br>",
      "<b>Treatment:</b> %{customdata}<br>",
      "<extra></extra>"
    )
  ) %>%
  layout(
    legend = list(
      orientation = "h",
      x = 0.5,
      y = -0.2,
      title = list(text = "<b>Treatment</b>", font = list(size = 13)),
      xanchor = "center",
      yanchor = "top"
    ),
    xaxis = list(title = list(
      text = "<b>Year of diagnosis</b>", font = list(size = 14)
    )),
    yaxis = list(title = list(
      text = "<b>Number of unique patients</b>", font = list(size = 14)
    ))
  )
```

# 3. Part 2: Dataset Construction

## 3.1. Extracted APACs

The number of APACs extracted was **`r n_apac_total`**. 

The ICD codes included in the analysis are presented in the table below. Observe that ICD code ended with ".x" indicates that last digit of the code was not recorded in the APAC.

```{r icd_table, echo = FALSE}
table_icd %>%
  auto_pack_rows(grouping_column = "CID_principal",
                 col.names = c("ICD", "Description", "N")) %>%
  kable_paper("hover", full_width = F)

# auto_pack_rows is a function from 'OncoDatasusR' package. For more information, see the documentation.
```

The following table presents the number of APACs and the corresponding number of CNSs before and after applying filter by procedure of non-interest. APACs without procedure codes are also excluded and appear in the table below as *'Missing Data'*. Note that in this analysis we chose to include all procedures.

```{r table_procedures, echo = FALSE}
rownames(table_procedures) <- c("APACs", "Unique codes", "Missing data")

table_procedures %>%
  kbl(col.names = c("Before", "After", "Diference")) %>%
  kable_paper("hover", full_width = F)
```

The table below shows the number of APACs by database (AQ and AR) and primary ICD (considering only the first three digits). This table is useful when the analysis involves more than one ICD.

```{r icd_sia_table, echo = FALSE}
table_icd_sia %>%
  kbl(col.names = c("ICD", "Description", "AQ", "AR", "Total")) %>%
  kable_paper("hover", full_width = F)  
```

## 3.2. Count of Patients by Unique CNS Codes

The Number of unique CNS codes identified was **`r n_before_deduplication`**

The table below presents the total number of CNS codes (variable AP_CNSPCN_ENC64) by procedure type (AQ or AR).Note that the total will be smaller than the sum of each dataset as the same CNS may have had APAC of AQ and AR.

```{r table_cns, echo = FALSE}
table_cns_sia %>%
  kbl(col.names = c("Database", "Unique codes")) %>%
  kable_paper("hover", full_width = F)
```

The table below presents some statistics to evaluate completeness, errors and inconsistencies in the dataset.

```{r data_quality, echo = FALSE}
table_data_quality %>%
  kbl(
    align = "c",
    col.names = c(
      "Missing data",
      "Incorrect data",
      "Inconsistent date",
      "Date previous to 1998",
      "Future date",
      "Total",
      "Proportion (%)"
    )
  ) %>%
  kable_paper("hover", full_width = F)
```

The table below presents the number of APACs according to the year of diagnosis, along with their respective frequencies, including incongruous values.

```{r table_year, echo = FALSE}
table_year %>%
  datatable(
    rownames = FALSE,
    colnames = c("Year", "Year of diagnosis", "Year of attendance"),
    options = list(
      pageLength = 6,
      dom = "tlpi",
      columnDefs = list(list(
        className = "dt-center", targets = 0:2
      ))
    )
  )
```

A total of **`r total_dt_error`** CNS codes were found with year of diagnosis prior to 1998 (10 years before de first year of interest). Using the rule of second smallest diagnosing date, it was possible to recover coherent dates for **`r total_replaced_diag_dt`** CNS codes.

The table below presents the number of procedures performed in the APACs by ICD code, considering all procedures performed in the period of analysis.

```{r table_procedures_performed, echo = FALSE}	
columns_to_center <- c(0, 2:(ncol(table_proced_total) - 1))

table_proced_total %>%
  datatable(
    rownames = FALSE,
    colnames = c("Procedure", "Description", colnames(table_proced_total)[3]),
    options = list(
      pageLength = 10,
      dom = "tlpi",
      columnDefs = list(list(
        className = "dt-center", targets = columns_to_center
      ))
    )
  )
```

The chart below shows the number of APACs by CNS code. Hover over the bars to see the detailed results.

```{r plot_cns, fig.width=8, echo = FALSE}
ggplotly(plot_cns_pcn, tooltip = c("x", "y"))
```

The table below is another way to look for the number of APACs by CNS code.

```{r table_apac_freq, echo = FALSE}
table_apac_cns %>%
  datatable(
    rownames = FALSE,
    colnames = c("Number of APACs", "Number of Patients (CNS)"),
    options = list(
      pageLength = 10,
      dom = "tlpi",
      columnDefs = list(list(
        className = "dt-center", targets = 0:1
      ))
    )
  )
```

The table below provides the number and proportion of CNS with different diagnosis dates, as well as the number of APACs associated with these CNS codes.

```{r table_incons_date, echo = FALSE}
table_inconsist_date %>%
  kbl(
    col.names = c(
      "Different diagnosis dates",
      "Patients (CNS)",
      "APAC",
      "Patients (CNS)",
      "APAC"
    ),
    align = "ccccc"
  ) %>%
  add_header_above(c(
    ` ` = 1,
    `Nº of records` = 2,
    `Proportion (%)` = 2
  )) %>%
  kable_paper("hover", full_width = F)
```

## 3.3. Analysis of Potential Duplicates

Assuming that a patient may have more than one CNS number, this stage involves searching for CNS codes that share the same information across key variables: first five digits of the postal code (ZIP5), ICD code, sex, diagnosis date, and 
year of birth.

The table below presents the number of missing values for the key variables. CNS codes with missing values are not eligible for this step of checking for duplicates and are kept as unique patients in the final dataset.

```{r table_key_vars, echo = FALSE}
table_key_var %>%
  kbl(col.names = c("Sex", "ZIP5", "Calculated birth year"),
      align = "c") %>%
  kable_paper("hover", full_width = F)
```

The table bellow shows the estimated number of patients by number of different CNS codes.

```{r table_duplicates, echo = FALSE}	
table_dup_cns %>%
  kbl(col.names = c("Number of CNS", "Patients", "Proportion"),
      align = "c") %>%
  kable_paper("hover", full_width = F)
```

* Number of unique codes after the duplicate analysis (considered as unique patients): **`r n_after_deduplication`**
* Absolute reduction in the number of unique codes: **`r n_before_deduplication - n_after_deduplication` records**
* Relative reduction compared to the dataset size: **`r var_linkage`%**

### 3.3.3. Final Adjustments

Patients excluded due to:

* Diagnosis year before 2008: **`r n_before_min_year`**
* Diagnosis year after 2022: **`r n_after_max_year`** 
* Procedure date before the diagnosis date: **`r dt_attend_before`**
* Age below minimum (under 18 years): **`r n_lt_min_age`**
* Age above maximum (over 99 years): **`r n_gt_max_age`**
* Removed due to selected sex: **`r n_rm_sex`**

Thus:

New patients counted during the study period: **`r n_patients_final`**
